### 2015-04-07 transactions.tex
Formatting edits
cf to rcf. Better formatting for operation names.

\vspace{-1em}
\begin{smathpar}
\forall a,b,c.~\txn{a}{b,c} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a}
\Rightarrow \vis{c}{a}
\end{smathpar}

@@

In order to illustrate the utility of declarative reasoning for transactions,
consider an extension of our running example to use two accounts (objects) --
current ($c$) and savings ($s$). Each account provides operations
\cf{withdraw}, \cf{deposit} and \cf{getBalance}, with the same contracts as
defined previously. We consider two transactions -- \cf{save(amt)}, which
transfers \cf{amt} from current to savings, and \cf{totalBalance}, which
returns the sum of the balances of individual accounts. Our goal is to ensure
that \cf{totalBalance} returns the result obtained from a consistent snapshot
of the object states. The \name code for the transactions is given below:
\vspace{-1em}

@@

While making both transactions serializable would ensure correctness,
distributed stores rarely offer serializable transactions since it is
unavailable and hinders scalability~\cite{BailisHAT}. As we will see, these
transactions can be satisfied with much weaker isolation guarantees. Despite
the atomicity offered by the transaction, anomalies are still possible. For
example, the two \cf{getBalance} operations in \cf{totalBalance} transactions
might be served by different replicas with distinct set of committed \cf{save}
transactions. If the first(second) \cf{getBalance} operation witness a
\cf{save} transaction that is not witnessed by the second(first)
\cf{getBalance} operation, then the balance returned will be less(greater) than
the actual balance. It is not immediately apparent which weakest isolation
guarantee will be sufficient to prevent the anomaly.

Instead, \name requires the programmer to simply state the consistency
requirement as a contract. Since we would like both the \cf{getBalance}
operations to witness the same set of \cf{save} transactions, we define the
constraint on \cf{totalBalance} transaction $\psi_{tb}$ as:

\vspace{-1em}
\begin{smathpar}
\begin{array}{l}
\cv_{tb} = \forall a:\rcf{getBalance}, b:\rcf{getBalance}, \\
\qquad (c:\rcf{withdraw} \vee \rcf{deposit}), (d:\rcf{withdraw} \vee \rcf{deposit}). \\
\qquad \txn{a,b}{c,d} ~\wedge~ \vis{c}{a} ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
\end{array}
\end{smathpar}

@@

\begin{figure}
\vspace{-2.3mm}
\centering
\subfigure[Atomicity]{
  \label{fig:txn_atomicity}
  \includegraphics[width=0.25\columnwidth]{Figures/TxnAtomic}
}
\hfill
\subfigure[Monotonic Atomic View]{
  \label{fig:txn_mav}
  \includegraphics[width=0.25\columnwidth] {Figures/TxnMAV}
}
\hfill
\subfigure[Repeatable Read]{
  \label{fig:txn_rr}
  \includegraphics[width=0.25\columnwidth]{Figures/TxnRR}
}
\caption{Semantics of transaction contracts. $x$ and $y$ are distinct objects.
The dotted line represents the visibility requested by the contracts.}
\label{fig:transaction}
\end{figure}

@@

\vspace{-1em}
\begin{smathpar}
\begin{array}{l}
\rcc = \forall a,b,c.~\txn{a}{b,c} \wedge \sameobj{b}{c} \\
\qquad \qquad \qquad \wedge~ \vis{b}{a} \Rightarrow \vis{c}{a}
\end{array}
\end{smathpar}

@@

\vspace{-0.6em}
\begin{smathpar}
\begin{array}{l}
\mavc = \forall a,b,c,d.~\txn{a,b}{c,d} ~\wedge~ \so{a}{b} ~\wedge~ \vis{c}{a} \\
\qquad\qquad\qquad\qquad ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
\end{array}
\end{smathpar}

@@

\vspace{-0.6em}
\begin{smathpar}
\begin{array}{l}
\rrc = \forall a,b,c,d.~\txn{a,b}{c,d} ~\wedge~ \vis{c}{a} \\
\qquad\qquad\qquad\quad ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
\end{array}
\end{smathpar}

>>>>

\begin{cmathpar}
\forall a,b,c.~\txn{a}{b,c} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a}
\Rightarrow \vis{c}{a}
\end{cmathpar}

@@

In order to illustrate the utility of declarative reasoning for transactions,
consider an extension of our running example to use two accounts (objects) --
current ($c$) and savings ($s$). Each account provides operations
\rcf{withdraw}, \rcf{deposit} and \rcf{getBalance}, with the same contracts as
defined previously. We consider two transactions -- \cf{save(amt)}, which
transfers \cf{amt} from current to savings, and \cf{totalBalance}, which
returns the sum of the balances of individual accounts. Our goal is to ensure
that \cf{totalBalance} returns the result obtained from a consistent snapshot
of the object states. The \name code for the transactions is given below:

@@

While making both transactions serializable would ensure correctness,
distributed stores rarely offer serializable transactions since it is
unavailable and hinders scalability~\cite{BailisHAT}. As we will see, these
transactions can be satisfied with much weaker isolation guarantees. Despite
the atomicity offered by the transaction, anomalies are still possible. For
example, the two \rcf{getBalance} operations in \cf{totalBalance} transactions
might be served by different replicas with distinct set of committed \cf{save}
transactions. If the first(second) \rcf{getBalance} operation witness a
\cf{save} transaction that is not witnessed by the second(first)
\rcf{getBalance} operation, then the balance returned will be less(greater) than
the actual balance. It is not immediately apparent which weakest isolation
guarantee will be sufficient to prevent the anomaly.

Instead, \name requires the programmer to simply state the consistency
requirement as a contract. Since we would like both the \rcf{getBalance}
operations to witness the same set of \cf{save} transactions, we define the
constraint on \cf{totalBalance} transaction $\psi_{tb}$ as:

\begin{cmathpar}
\begin{array}{l}
\cv_{tb} = \forall a:\rcf{getBalance}, b:\rcf{getBalance}, \\
\qquad (c:\rcf{withdraw} \vee \rcf{deposit}), (d:\rcf{withdraw} \vee \rcf{deposit}). \\
\qquad \txn{a,b}{c,d} ~\wedge~ \vis{c}{a} ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
\end{array}
\end{cmathpar}

@@

\begin{figure}
\centering
\subfigure[Atomicity]{
  \label{fig:txn_atomicity}
  \includegraphics[width=0.25\columnwidth]{Figures/TxnAtomic}
}
\hfill
\subfigure[Monotonic Atomic View]{
  \label{fig:txn_mav}
  \includegraphics[width=0.25\columnwidth] {Figures/TxnMAV}
}
\hfill
\subfigure[Repeatable Read]{
  \label{fig:txn_rr}
  \includegraphics[width=0.25\columnwidth]{Figures/TxnRR}
}
\caption{Semantics of transaction contracts. $x$ and $y$ are distinct objects.
The dotted line represents the visibility requested by the contracts.}
\label{fig:transaction}
\end{figure}

@@

\begin{cmathpar}
\begin{array}{l}
\rcc = \forall a,b,c.~\txn{a}{b,c} \wedge \sameobj{b}{c} \\
\qquad \qquad \qquad \wedge~ \vis{b}{a} \Rightarrow \vis{c}{a}
\end{array}
\end{cmathpar}

@@

\begin{cmathpar}
\begin{array}{l}
\mavc = \forall a,b,c,d.~\txn{a,b}{c,d} ~\wedge~ \so{a}{b} ~\wedge~ \vis{c}{a} \\
\qquad\qquad\qquad\qquad ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
\end{array}
\end{cmathpar}

@@

\begin{cmathpar}
\begin{array}{l}
\rrc = \forall a,b,c,d.~\txn{a,b}{c,d} ~\wedge~ \vis{c}{a} \\
\qquad\qquad\qquad\quad ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
\end{array}
\end{cmathpar}
